Using Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2020-12-10 thesis/thesis.tex
thesis: Update State of the Art

% TODO: Hablar de KMedias y DEC

@@

>>>>

@@

\section{Clustering Algorithms}

As we will explain later, what we need to accomplish our aim is an algorithm that manages to make data groups based on the dynamic properties of the stars.

There are several clustering algorithms: KMeans, Mean-Shift Clustering, DBSCAN \cite{ester1996density} among other.
While each one behaves better according ot the distribution of the objects to clusterize, we have choosen KMeans by its simplicity and good results.

KMeans requires a single parameter, the number of clusters to build (N).
The algorithm looks for N center points which are vectors of the same dimension as the number of selected features.
KMeans starts by making an initial groups configuration and then, it reassigns objects to other groups iteratively by minimizing
the distance among points inside the new group and maximizing the distance with the centers of the other groups.

The algorithm stops once a maximum number of iterations is reached, or when changes between iterations is lower than a minimum.

This algorithm works well and gives good results at first approach.
However, we need to set a large number of clusters to find the open cluster we are looking for.
This fact complicates the identification of the OC and many times the OC contains to many outliers.
For that reason, we have search for a KMeans refinement based on an Artificial Neural Network (ANN).

The \emph{Unsupervised Deep Embedding for Clustering Analysis} model or \emph{DEC} \cite{xie2016unsupervised} takes KMeans as its starting point,
but then, it trains an autoencoder to reduce the feature space and pass this transformed data through a Clustering Layer which refines the previous selection.

The mechanism of this model is explained in Section \ref{sec:deep_embedding_clustering} and its implementation is available in \verb|cdalvaro.ml.dec| package.
